---
title: "P8105 HW6"
author: Shihui Zhu sz3029
output: github_document
---

```{r setup, include=FALSE}
# Reproducibility
set.seed(1)

# This chunk loads all the packages used in this homework
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)
library(leaps)

# General figure set up
knitr::opts_chunk$set(
  # display the code in github doc
  echo = TRUE,
  # hide warning messages
  warning = FALSE,
  message = FALSE,
  # set the figure to be 8 x 6, and the proportion it takes to be 95%
  fig.width = 10,
  fig.height = 8, 
  out.width = "90%"
)

# setting a global options for continuous data color family and a different format to set discrete data to have a color family
options(
  ggplot2.countinuous.colour = "viridis",
  ggplot2.countinuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# have a minimal theme and legends at the bottom
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

### Import data

Load and clean the data for regression analysis
```{r import_data, message=FALSE}
birthweight_df <- 
  read_csv("data/birthweight.csv") %>%
  drop_na() %>%
  mutate(
    # change the numbers to text for a better readability
    babysex = ifelse(babysex == 1, 'male', 'female'),
    # convert to factor
    babysex = fct_infreq(babysex),
    # change the numbers to text for a better readability
    frace = case_when(
      frace == 1 ~ 'white',
      frace == 2 ~ 'black',
      frace == 3 ~ 'asian',
      frace == 4 ~ 'puerto_rican',
      frace == 8 ~ 'other',
      frace == 9 ~ 'unknown'),
    # convert to factor
    frace = fct_infreq(frace),
    # change the numbers to text for a better readability
    malform = ifelse(malform == 1, 'present', 'absent'),
    # convert to factor
    malform = fct_infreq(malform),
    # change the numbers to text for a better readability
    mrace = case_when(
      mrace == 1 ~ 'white',
      mrace == 2 ~ 'black',
      mrace == 3 ~ 'asian',
      mrace == 4 ~ 'puerto_rican',
      mrace == 8 ~ 'other'),
    # convert to factor
    mrace = fct_infreq(mrace),
    )

birthweight_df
```

### Build model

They are 19 variables in total. We want to select the most important variables that affect baby's birth weight to build the regression model. We will use a data-driven model building process, backwards elimination.

1. Select variables using backwards elimination
```{r mlr}
mult.fit = lm(bwt ~ ., data = birthweight_df)
model.best = step(mult.fit, direction='backward')
```

The "best" model proposed by backward selection is:

```{r best}
model.best$coefficients %>% knitr::kable(digits = 3, col.names = c("Coefficient"))
```

2. Select the variables and build the MLR model

```{r model}
model_1 = lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + fincome + 
               mheight + mrace + parity +
               ppwt + smoken, data = birthweight_df)

model_1 %>%
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^babysex", "Baby sex: "),
         term = str_replace(term, "^mrace", "Mother race: ")) %>% 
  knitr::kable(digits = 3)
```

### Diagnostics

```{r resid}
modelr::add_residuals(birthweight_df, model_1) %>% head(10) %>% knitr::kable(digits = 3)

modelr::add_predictions(birthweight_df, model_1) %>% head(10) %>% knitr::kable(digits = 3)
```

Show an example model residuals against fitted values

```{r plot_ex}
birthweight_df %>% 
  modelr::add_residuals(model_1) %>% 
  ggplot(aes(x = gaweeks, y = resid)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Residuals v.s. Fitted Values of Birth Weight By Gestational Age in Weeks",
    x = 'Gestational Age in Weeks',
    y = 'Baby\'s Weight'
  )
```

## Compare the model to two others

Use `cross_mc` to prepare CV training and testing sets
```{r}
#Splitting
cv_df = 
  modelr::crossv_mc(birthweight_df, length(birthweight_df)) 
#training
cv_df %>% pull(train) %>% .[[1]] %>% as_tibble
#testing
cv_df %>% pull(test) %>% .[[1]] %>% as_tibble

```

Compare the above model to two others

* One using length at birth and gestational age as predictors (main effects only)

* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
```{r}
cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

model_1 = lm(bwt ~ babysex + bhead + blength + frace + gaweeks + 
               mheight + momage + mrace +
               ppwt + smoken + wtgain, data = birthweight_df)

cv_df = 
  cv_df %>% 
  mutate(
    mod_1  = map(train, ~lm(bwt ~ babysex + bhead + blength + frace + gaweeks + 
               mheight + momage + mrace +
               ppwt + smoken + wtgain, data = .x)),
    mod_2  = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    mod_3  = map(train, ~lm(bwt ~ babysex + bhead + blength + babysex * bhead * blength, data = .x))) %>% 
  mutate(
    rmse_mod1 = map2_dbl(mod_1, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_mod2 = map2_dbl(mod_2, test, ~modelr::rmse(model = .x, data = .y)),
    rmse_mod3 = map2_dbl(mod_3, test, ~modelr::rmse(model = .x, data = .y)))

cv_df %>% select(.id, rmse_mod1:rmse_mod3) %>% knitr::kable(digits = 3)
```

## Problem 2

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

