---
title: "P8105 HW6"
author: Shihui Zhu sz3029
output: github_document
---

```{r setup, include=FALSE}
# Reproducibility
set.seed(1)

# This chunk loads all the packages used in this homework
library(tidyverse)
library(viridis)
library(ggridges)
library(patchwork)

library(readxl)

# General figure set up
knitr::opts_chunk$set(
  # display the code in github doc
  echo = TRUE,
  # hide warning messages
  warning = FALSE,
  message = FALSE,
  # set the figure to be 8 x 6, and the proportion it takes to be 95%
  fig.width = 10,
  fig.height = 8, 
  out.width = "90%"
)

# setting a global options for continuous data color family and a different format to set discrete data to have a color family
options(
  ggplot2.countinuous.colour = "viridis",
  ggplot2.countinuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# have a minimal theme and legends at the bottom
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

### Import data

Load and clean the data for regression analysis
```{r import_data, message=FALSE}
birthweight_df <- 
  read_csv("data/birthweight.csv") %>%
  drop_na() %>%
  mutate(
    # change the numbers to text for a better readability
    babysex = ifelse(babysex == 1, 'male', 'female'),
    # convert to factor
    babysex = fct_infreq(babysex),
    # change the numbers to text for a better readability
    frace = case_when(
      frace == 1 ~ 'white',
      frace == 2 ~ 'black',
      frace == 3 ~ 'asian',
      frace == 4 ~ 'puerto_rican',
      frace == 8 ~ 'other',
      frace == 9 ~ 'unknown'),
    # convert to factor
    frace = fct_infreq(frace),
    # change the numbers to text for a better readability
    malform = ifelse(malform == 1, 'present', 'absent'),
    # convert to factor
    malform = fct_infreq(malform),
    # change the numbers to text for a better readability
    mrace = case_when(
      mrace == 1 ~ 'white',
      mrace == 2 ~ 'black',
      mrace == 3 ~ 'asian',
      mrace == 4 ~ 'puerto_rican',
      mrace == 8 ~ 'other'),
    # convert to factor
    mrace = fct_infreq(mrace),
    
    )
```

### Build model

They are 19 variables in total. We want to select the most important variables that affect baby's birth weight. 

1. Compute the marginal pmf/pdf of all variables, and get the p-values
```{r mlr}
# pmf/pdf
compute_pmf <- function(x) {
  fit = lm(bwt ~ x, data = birthweight_df)
  p_value = summary(fit)$coefficients[,4][2]
  return(p_value)
}

p_values = birthweight_df %>%
  select(-bwt) %>%
  map_dbl(compute_pmf)

p_values = bind_rows(p_values)

p_values %>% pivot_longer(
  everything(),
  values_to = 'p_value',
  names_to = 'variables') %>%
  mutate(
    significant = ifelse(p_value > 0.025 | is.na(p_value), 'No', 'Yes')
  ) %>%
  knitr::kable(digits = 3)
```

```{r plots}
# graph
i = 1
compute_plist <- function(x) {
  if (is.factor(x)) {
    pl = birthweight_df %>% 
    ggplot(aes(x = x, y = bwt)) +
    geom_boxplot() +
    labs(
      x = '',
      y = "birth weight",
    )
  } else {
    pl = birthweight_df %>% 
    ggplot(aes(x = x, y = bwt)) +
    geom_point() +
    geom_smooth(method = "lm") +
    labs(
      x = '',
      y = "birth weight",
    )
  }
  i = i + 1
  return(pl)
}

plist = birthweight_df %>%
  select(-bwt) %>%
  map(compute_plist)

labels = birthweight_df %>%
  select(-bwt) %>%
  names()

for (i in 1:length(plist)) {
  plist[[i]] = plist[[i]] + labs(x = labels[[i]])
}

wrap_plots(plist, ncol = 4)
```

p-value

graph

ppwt and delwt seems to be redundant with wtgain, delete delwt

2. Select the variables and build the MLR model

```{r model}
model_1 = lm(bwt ~ babysex + bhead + blength + frace + gaweeks + 
               mheight + momage + mrace +
               ppwt + smoken + wtgain, data = birthweight_df)

model_1 %>%
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^babysex", "Baby sex: "),
         term = str_replace(term, "^frace", "Father race: "),
         term = str_replace(term, "^mrace", "Mother race: ")) %>% 
  knitr::kable(digits = 3)
```

### Diagnostics

```{r resid}
modelr::add_residuals(birthweight_df, model_1) %>% head(10) %>% knitr::kable(digits = 3)

modelr::add_predictions(birthweight_df, model_1) %>% head(10) %>% knitr::kable(digits = 3)
```

Show an example model residuals against fitted values

```{r plot_ex}
birthweight_df %>% 
  modelr::add_residuals(model_1) %>% 
  ggplot(aes(x = mrace, y = resid)) + 
  geom_violin() +
  labs(
    title = "Violin Residuals Plots By Category of Mother's Race",
    x = 'Mother\'s race',
    y = 'Baby\'s Weight'
  )
```


